---
title: 'All the R scripts for publication_children'
author: 'Jyoti Dalal'
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(message=F, echo = F, warning = F, dpi = 300)

```


```{r, include=FALSE}
rm(list = ls())
library(pacman) # loading all the required packages

p_load(char=c("lattice","ggplot2","highr", "stringr", "markdown", 
              "knitr", "mime", "rmarkdown","reactable", "reshape",
              "tidyverse", "ggrepel", "lubridate", "sf","incidence",
              "magrittr", "ggspatial", "RColorBrewer", "gridExtra","zoo","readxl"))
```


```{r, include=FALSE}
# com_data = NULL
# lines <- list()
# countries_csv <- list.files("~/publication-children/data") # listing all data files
# countries_names <- str_trim(str_remove(countries_csv, "_clean.csv"), "both") # extracting country names from data files
# 
# calculating_case_deaths = function(data, gender) # to compute confirmed cases + 
#   #deaths in various age_groups
# {
#   return(data %>% filter(patinfo_sex == gender) %>%
#            group_by(country, "gender" = patinfo_sex, "age groups" = age_cat) %>% 
#            summarise(confirmed_cases_each_group = n(), 
#                      confirmed_deaths_each_group = sum(death, na.rm = T))
#   )
# }
# 
# fetching_data = function(df, country)
# {
#   return(mutate(df, country = str_to_sentence(country), patinfo_ageonset_years = as.double(patinfo_ageonset_years),
#                 #patinfo_ID = as.character(patinfo_ID), 
#                 patcourse_datedeath = as.Date(patcourse_datedeath), report_classif = toupper(as.character(report_classif)), patcourse_status = toupper(as.character(patcourse_status))) %>%
#            mutate(report_date = if_else(is.na(report_date), as.Date(Lab_resdate), as.Date(report_date) )) %>%
#            mutate(report_date = if_else(is.na(report_date), as.Date(consultation_dateHF), as.Date(report_date) )) %>%
#            mutate(report_date = if_else(is.na(report_date), as.Date(Lab_datetaken), as.Date(report_date) )) %>%
#            mutate(report_date = if_else(is.na(report_date), as.Date(patcourse_dateonset), as.Date(report_date) )) %>%
#            mutate(report_date = if_else(is.na(report_date), as.Date(patcourse_datedeath), as.Date(report_date) )) %>%
#            mutate(report_date = if_else(is.na(report_date), as.Date(patcourse_datedischarge), as.Date(report_date) ))%>%
#            mutate(patcourse_status = if_else((patcourse_status=='ALIVE' | patcourse_status== 'RECOVERED') & !is.na(patcourse_datedeath), 'DEAD', patcourse_status)) %>%
#            mutate(patcourse_status = if_else(is.na(patcourse_status) & !is.na(patcourse_datedeath), 'DEAD', patcourse_status)) %>%
#            mutate(patcourse_status = if_else(is.na(patcourse_status) & !is.na(patcourse_datedischarge), 'RECOVERED', patcourse_status)) %>%
#            mutate(report_classif = if_else(is.na(report_classif) & (sum(is.na(report_classif))==nrow(df)), 'CONFIRMED', report_classif)) %>%
#            mutate(death = if_else(patcourse_status=='DEAD', 1, 0))%>%
#            mutate(
#              age_cat = cut(patinfo_ageonset_years, c(-1,4,11,17, 150), 
#                            labels=c("0-4", "5-11","12-17", "17+"))) %>%
#            filter(report_date <=lubridate::ymd("2020-09-01") & report_date >=lubridate::ymd("2020-04-01"))%>%
#            filter(report_classif=='CONFIRMED') %>% 
#            filter(!is.na(patinfo_sex))) %>%
#     select(country, report_date, patinfo_sex, patinfo_ageonset_years, death, 
#            age_cat, report_classif, patcourse_status)
# }
# # Reading the data set
# lines <- list()
# for (country in countries_names){
#   print(country)
#   lines[country] <- readLines(paste("~/publication-children/data/",country, "_clean.csv", sep =''), n = 1)
#   #lines[country] <- readLines(here("child_pub/data_analy", glue('{country}_clean.csv')), n = 1)
#   
#   #******* reading the CSVs *********#
#   
#   if(grepl(',', lines[country])){
#     print(country)
#     print("1")
#     data1 <- as.data.frame(read.csv(paste("~/publication-children/data/", country, "_clean.csv", sep ='')),na.strings =
#                              ".") # , separated
#     print(country)}
#   else {
#     print(country)
#     print("2")
#     data1 <- as.data.frame(read.csv2(paste("~/publication-children/data/", country, "_clean.csv", sep ='')), na.strings =
#                              ".") #} # , separated
#     print(country)}
#   
#   data = fetching_data(data1, country) # only CONFIRMED cases
#   
#   #****** CONFIRMED cases + deaths in various age-groups ********#
#   
#   m_data = calculating_case_deaths(data, 'M') # in males
#   f_data = calculating_case_deaths(data, 'F') # in females
#   com_data = rbind.data.frame(com_data, m_data, f_data) 
# }
```
 
```{r, include=FALSE}

df_com = NULL
lines <- list()
countries_csv <- list.files("~/publication-children/data") # listing all data files
countries_names <- str_trim(str_remove(countries_csv, "_clean.csv"), "both") # extracting country names from data files

fetching_data = function(df, country)
{
  return(mutate(df, country = str_to_sentence(country), patinfo_ageonset_years = as.double(patinfo_ageonset_years),
                patcourse_datedeath = as.Date(patcourse_datedeath), report_classif = toupper(as.character(report_classif)),  
                patcourse_status = toupper(as.character(patcourse_status))) %>%
           mutate(report_date = if_else(is.na(report_date), as.Date(Lab_resdate), as.Date(report_date) )) %>%
           mutate(report_date = if_else(is.na(report_date), as.Date(consultation_dateHF), as.Date(report_date) )) %>%
           mutate(report_date = if_else(is.na(report_date), as.Date(Lab_datetaken), as.Date(report_date) )) %>%
           mutate(report_date = if_else(is.na(report_date), as.Date(patcourse_dateonset), as.Date(report_date) )) %>%
           mutate(report_date = if_else(is.na(report_date), as.Date(patcourse_datedeath), as.Date(report_date) )) %>%
           mutate(report_date = if_else(is.na(report_date), as.Date(patcourse_datedischarge), as.Date(report_date) ))%>%
           mutate(patcourse_status = if_else((patcourse_status=='ALIVE' | patcourse_status== 'RECOVERED') & !is.na(patcourse_datedeath), 'DEAD', patcourse_status)) %>%
           mutate(patcourse_status = if_else(is.na(patcourse_status) & !is.na(patcourse_datedeath), 'DEAD', patcourse_status)) %>%
           mutate(patcourse_status = if_else(is.na(patcourse_status) & !is.na(patcourse_datedischarge), 'RECOVERED', patcourse_status)) %>%
           mutate(report_classif = if_else(is.na(report_classif) & (sum(is.na(report_classif))==nrow(df)), 'CONFIRMED', report_classif)) %>%
           mutate(death = if_else(patcourse_status=='DEAD', 1, 0))%>%
           mutate(
             age_cat = cut(patinfo_ageonset_years, c(-1,4,11,17, 150), 
                           labels=c("0-4", "5-11","12-17", "17+"))) %>%
           filter(report_date <=lubridate::ymd("2020-09-01") & report_date >=lubridate::ymd("2020-04-01"))%>%
           filter(report_classif=='CONFIRMED') %>% filter(!is.na(patinfo_ageonset_years)) %>% filter(!is.na(patinfo_sex)))  %>%
    select(country, report_date, patinfo_sex, patinfo_ageonset_years, death, 
           age_cat, report_classif, patinfo_sex, patcourse_status)
}
# Reading the data set
lines <- list()
for (country in countries_names){
  print(country)
  lines[country] <- readLines(paste("~/publication-children/data/",country, "_clean.csv", sep =''), n = 1)
  #lines[country] <- readLines(here("child_pub/data_analy", glue('{country}_clean.csv')), n = 1)
  
  #******* reading the CSVs *********#
  
  if(grepl(',', lines[country])){
    print(country)
    print("1")
    data1 <- as.data.frame(read.csv(paste("~/publication-children/data/", country, "_clean.csv", sep ='')),na.strings =
                             ".") # , separated
    print(country)}
  else {
    print(country)
    print("2")
    data1 <- as.data.frame(read.csv2(paste("~/publication-children/data/", country, "_clean.csv", sep ='')), na.strings =
                             ".") #} # , separated
    print(country)}
  
  data = fetching_data(data1, country) # only CONFIRMED cases
  
  data_f = bind_rows(df_com, data)
  df_com = data_f
  }

#missing_gender = sum(is.na(data$patinfo_sex)) #***** checking for missing gender info ******#
#missing_status = sum(is.na(data$patcourse_status)) #***** checking for missing patient's status info ******#
  #missing_age = sum(is.na(data$patinfo_ageonset_years)) #***** checking for missing patient's age info ******#
```

## Table 1 (sample characteristics). ##

**number of total cases** 
In each strata, total number should be equal to = 51964

```{r}
nrow(df_com)
```

**Sex**
  
```{r}
table(df_com$patinfo_sex)

round(100*(table(df_com$patinfo_sex)/nrow(df_com)), 1) # in percentages
```

**Missing Value in Sex**
  
```{r}
length(which(is.na(df_com$patinfo_sex)))
```

**Age**
  
```{r}
table(df_com$age_cat)
round(100*(table(df_com$age_cat)/nrow(df_com)), 1)
```

**Missing Value in Age**
  
```{r}
length(which(is.na(df_com$age_cat)))
```

**Clinical outcome**
  
```{r}
table(df_com$patcourse_status)
```

**Missing Value in Clinical outcome**
I have counted '8' values with UNKNOWN status in NAs.

```{r}
(length(which(is.na(df_com$patcourse_status))) + 8)
round(100*(length(which(is.na(df_com$patcourse_status))) + 8)/nrow(df_com), 1) # fraction of NAs
```

ALIVE, DEAD, RECOVERED (%)

```{r}
round(100*(table(df_com$patcourse_status)/nrow(df_com)), 1)
```

**Country wise **
  
```{r}
table(df_com$country)
round(100*table(df_com$country)/nrow(df_com), 1)
```

```{r, include=FALSE}
com_data = NULL
lines <- list()
countries_csv <- list.files("~/publication-children/data") # listing all data files
countries_names <- str_trim(str_remove(countries_csv, "_clean.csv"), "both") # extracting country names from data files

calculating_case_deaths = function(data) # to compute confirmed cases +
  #deaths in various age_groups
{
  return(data %>%
           group_by(country, "age groups" = age_cat) %>%
           summarise(confirmed_cases_each_group = n(),
                     confirmed_deaths_each_group = sum(death, na.rm = T))
  )
}

fetching_data = function(df, country)
{
  return(mutate(df, country = str_to_sentence(country), patinfo_ageonset_years = as.double(patinfo_ageonset_years), patcourse_datedeath = as.Date(patcourse_datedeath), report_classif = toupper(as.character(report_classif)), patcourse_status = toupper(as.character(patcourse_status))) %>%
           mutate(report_date = if_else(is.na(report_date), as.Date(Lab_resdate), as.Date(report_date) )) %>%
           mutate(report_date = if_else(is.na(report_date), as.Date(consultation_dateHF), as.Date(report_date) )) %>%
           mutate(report_date = if_else(is.na(report_date), as.Date(Lab_datetaken), as.Date(report_date) )) %>%
           mutate(report_date = if_else(is.na(report_date), as.Date(patcourse_dateonset), as.Date(report_date) )) %>%
           mutate(report_date = if_else(is.na(report_date), as.Date(patcourse_datedeath), as.Date(report_date) )) %>%
           mutate(report_date = if_else(is.na(report_date), as.Date(patcourse_datedischarge), as.Date(report_date) ))%>%
           mutate(patcourse_status = if_else((patcourse_status=='ALIVE' | patcourse_status== 'RECOVERED') & !is.na(patcourse_datedeath), 'DEAD', patcourse_status)) %>%
           mutate(patcourse_status = if_else(is.na(patcourse_status) & !is.na(patcourse_datedeath), 'DEAD', patcourse_status)) %>%
           mutate(patcourse_status = if_else(is.na(patcourse_status) & !is.na(patcourse_datedischarge), 'RECOVERED', patcourse_status)) %>%
           mutate(report_classif = if_else(is.na(report_classif) & (sum(is.na(report_classif))==nrow(df)), 'CONFIRMED', report_classif)) %>%
           mutate(death = if_else(patcourse_status=='DEAD', 1, 0))%>%
           mutate(
             age_cat = cut(patinfo_ageonset_years, c(-1,4,11,17, 150),
                           labels=c("0-4", "5-11","12-17", "17+"))) %>%
           filter(report_date <=lubridate::ymd("2020-09-01") & report_date >=lubridate::ymd("2020-04-01"))%>%
           filter(report_classif=='CONFIRMED') %>%
           filter(!is.na(patinfo_sex)) %>% filter(!is.na(patinfo_ageonset_years))) %>%
    select(country, report_date, patinfo_sex, patinfo_ageonset_years, death,
           age_cat, report_classif, patcourse_status)
}
# Reading the data set
lines <- list()
for (country in countries_names){
  print(country)
  lines[country] <- readLines(paste("~/publication-children/data/",country, "_clean.csv", sep =''), n = 1)
  #lines[country] <- readLines(here("child_pub/data_analy", glue('{country}_clean.csv')), n = 1)

  #******* reading the CSVs *********#

  if(grepl(',', lines[country])){
    print(country)
    print("1")
    data1 <- as.data.frame(read.csv(paste("~/publication-children/data/", country, "_clean.csv", sep ='')),na.strings =
                             ".") # , separated
    print(country)}
  else {
    print(country)
    print("2")
    data1 <- as.data.frame(read.csv2(paste("~/publication-children/data/", country, "_clean.csv", sep ='')), na.strings =
                             ".") #} # , separated
    print(country)}

  data = fetching_data(data1, country) # only CONFIRMED cases

  #****** CONFIRMED cases + deaths in various age-groups ********#

  data_cd = calculating_case_deaths(data)
  com_data = rbind.data.frame(com_data, data_cd)
}
```

## Table 2. ##

```{r}
knitr::kable(com_data)
```

### Figure 1. Trends in 15 countries next to each other ###
Figure 1.

```{r}
# script to show the trends in 15 countries, jusr next to ech other
age_pop1 = as.data.frame(read.csv("~/publication-children/age_pop1.csv"))

fetching_data = function(df, country)
{
  return(mutate(df, country = str_to_sentence(country), patinfo_ageonset_years = as.double(patinfo_ageonset_years),
                #patinfo_ID = as.character(patinfo_ID),
                patcourse_datedeath = as.Date(patcourse_datedeath), report_classif = toupper(as.character(report_classif)), patcourse_status = toupper(as.character(patcourse_status))) %>%
           mutate(report_date = if_else(is.na(report_date), as.Date(Lab_resdate), as.Date(report_date) )) %>%
           mutate(report_date = if_else(is.na(report_date), as.Date(consultation_dateHF), as.Date(report_date) )) %>%
           mutate(report_date = if_else(is.na(report_date), as.Date(Lab_datetaken), as.Date(report_date) )) %>%
           mutate(report_date = if_else(is.na(report_date), as.Date(patcourse_dateonset), as.Date(report_date) )) %>%
           mutate(report_date = if_else(is.na(report_date), as.Date(patcourse_datedeath), as.Date(report_date) )) %>%
           mutate(report_date = if_else(is.na(report_date), as.Date(patcourse_datedischarge), as.Date(report_date) ))%>%
           mutate(patcourse_status = if_else((patcourse_status=='ALIVE' | patcourse_status== 'RECOVERED') & !is.na(patcourse_datedeath), 'DEAD', patcourse_status)) %>%
           mutate(patcourse_status = if_else(is.na(patcourse_status) & !is.na(patcourse_datedeath), 'DEAD', patcourse_status)) %>%
           mutate(patcourse_status = if_else(is.na(patcourse_status) & !is.na(patcourse_datedischarge), 'RECOVERED', patcourse_status)) %>%
           mutate(report_classif = if_else(is.na(report_classif) & (sum(is.na(report_classif))==nrow(df)), 'CONFIRMED', report_classif)) %>%
           mutate(death = if_else(patcourse_status=='DEAD', 1, 0))%>%
           mutate(
             age_cat = cut(patinfo_ageonset_years, c(-1,4,11,17, 150),
                           labels=c("0-4", "5-11","12-17", "17+"))) %>%
           filter(report_date<=lubridate::ymd("2020-09-01") & report_date>=lubridate::ymd("2020-04-01"))%>%
           filter(report_classif=='CONFIRMED') %>% filter(!is.na(patinfo_ageonset_years)) %>% filter(!is.na(patinfo_sex))) %>%
    select(country, report_date, patinfo_sex, patinfo_ageonset_years, death, age_cat, report_classif)
}

lines <- list()
countries_csv <- list.files("~/publication-children/data")
countries_names <- str_trim(str_remove(countries_csv, "_clean.csv"), "both")

#age_pop = as.data.frame(read.csv("~/publication-children/age_pop.csv"))

for(country in countries_names){   
#for(country in countries_names[grepl('Eswat', countries_names)]){
  print(country)
  lines[country] <- readLines(paste("~/publication-children/data/",country, "_clean.csv", sep =''), n = 1)
  
  if(grepl(',', lines[country])){
    print(country)
    print("1")
    data1 <- as.data.frame(read.csv(paste("~/publication-children/data/", country, "_clean.csv", sep ='')),na.strings =
                             ".") # , separated
    print(country)}
  else {
    print(country)
    print("2")
    data1 <- as.data.frame(read.csv2(paste("~/publication-children/data/", country, "_clean.csv", sep ='')), na.strings =
                             ".") 
    print(country)}
  #print()
  
  data = fetching_data(data1, country) # only CONFIRMED cases
  
  #colnames = X0.4"      "X5.11"     "X12.17"    "X17." 
  age_pop_1 = age_pop1$X0.4[age_pop1$Countries==country]
  age_pop_2 = age_pop1$X5.11[age_pop1$Countries==country]
  age_pop_3 = age_pop1$X12.17[age_pop1$Countries==country]
  
  pop1 = (100000)/(age_pop_1)  # 0-4,
  pop2 = (100000)/(age_pop_2) # 5-11
  pop3 =(100000)/(age_pop_3) # 12-17
  
  ## epi_curves for different age-categories:
  epi_per_age1 = data %>% group_by(age_cat, report_date) %>%
    summarise(Reported_this_day = n()) %>%
    filter(age_cat=='0-4') %>% 
    complete(report_date = seq(as.Date('2020-04-01'), as.Date('2020-09-01') , by = "day")) %>%
    mutate(Reported_this_day = ifelse(is.na(Reported_this_day), 0, Reported_this_day)) %>%
    mutate(cumulat_cases = cumsum(Reported_this_day)) %>% mutate(Reported_this_day_scaled1 = Reported_this_day*pop1) %>%
    mutate(cumulat_cases_scaled1 = cumulat_cases*pop1) %>%
    mutate(sevendayavg_cum1=rollmean(x = cumulat_cases_scaled1, k=7, align = "right",
                                     fill = NA )) %>%
    mutate(sevendayavg1=rollmean(x = Reported_this_day_scaled1, k=7, align = "right",
                                 fill = NA )) # 7-days rolling average of scaled-cases
  
  epi_per_age2 = data %>% group_by(age_cat, report_date) %>%
    summarise(Reported_this_day = n()) %>%
    filter(age_cat=='5-11') %>% complete(report_date = seq(as.Date('2020-04-01'),
                                                           as.Date('2020-09-01') , by = "day")) %>%
    mutate(Reported_this_day = ifelse(is.na(Reported_this_day), 0, Reported_this_day)) %>%
    mutate(cumulat_cases = cumsum(Reported_this_day)) %>%
    mutate(Reported_this_day_scaled2 = Reported_this_day*pop2) %>%
    mutate(cumulat_cases_scaled2 = cumulat_cases*pop2) %>%
    mutate(sevendayavg_cum2=rollmean(x = cumulat_cases_scaled2, k=7, align = "right",
                                     fill = NA )) %>%
    mutate(sevendayavg2=rollmean(x = Reported_this_day_scaled2, k=7, align = "right",
                                 fill = NA )) # 7-days rolling average of scaled-cases
  
  
  epi_per_age3 = data %>% group_by(age_cat, report_date) %>%
    summarise(Reported_this_day = n()) %>%
    filter(age_cat=='12-17') %>% complete(report_date = seq(as.Date('2020-04-01'),
                                                            as.Date('2020-09-01') , by = "day")) %>%
    mutate(Reported_this_day = ifelse(is.na(Reported_this_day), 0, Reported_this_day)) %>%
    mutate(cumulat_cases = cumsum(Reported_this_day)) %>%
    mutate(Reported_this_day_scaled3 = Reported_this_day*pop3) %>%
    mutate(cumulat_cases_scaled3 = cumulat_cases*pop3) %>%
    mutate(sevendayavg_cum3=rollmean(x = cumulat_cases_scaled3, k=7, align = "right",
                                     fill = NA )) %>%
    mutate(sevendayavg3=rollmean(x = Reported_this_day_scaled3, k=7, align = "right",
                                 fill = NA )) # 7-days rolling average of scaled-cases
  
  assign(paste("epi_1_", country, sep = ""), epi_per_age1)
  assign(paste("epi_2_", country, sep = ""), epi_per_age2)
  assign(paste("epi_3_", country, sep = ""), epi_per_age3)
  
  #print(country)
  
  comb_plots = epi_per_age1 %>% ggplot() +
    geom_line(aes(x=report_date, y=(sevendayavg1), col = "blue"), size = 0.8) +
    geom_line(aes(x=report_date, y=epi_per_age2$sevendayavg2, col = 'red'), size = 0.8, linetype = 'dashed')+
    geom_line(aes(x=report_date, y=epi_per_age3$sevendayavg3, col = 'brown'), size = 0.8, linetype = 'dotted')+
    labs(x=" ", y = " ") + 
    theme(axis.text.x=element_text(size=0.001, angle = 90)) +
    scale_color_identity(name = "Age groups",
                         breaks = c("blue", "red", 'brown'),
                         labels = c("0-4",'5-11','12-17'),
                         guide = "legend")  +
    theme_bw() +
    theme(legend.position = "top",
          legend.title = element_blank(),
          legend.text = element_text(size = 10))
  
  print(comb_plots)
  #write.csv(epi_per_age1, paste(country, "1_0_4.csv")) # 0-4
  #write.csv(epi_per_age2, paste(country, "2_5_11.csv")) # 5-11
  #write.csv(epi_per_age3, paste(country, "3_12_17.csv")) # 12-17
}

```

### Children vs adults (combining data from all the 15 countries) 
Figure 3.

```{r}
# script used to create children vs adults trendline, combining data from all the countries.

fetching_data_2 = function(df, country)
{
  return(mutate(df, country = str_to_sentence(country), patinfo_ageonset_years = as.double(patinfo_ageonset_years),
                #patinfo_ID = as.character(patinfo_ID), 
                patcourse_datedeath = as.Date(patcourse_datedeath),
                report_classif = toupper(as.character(report_classif)), patcourse_status = toupper(as.character(patcourse_status))) %>% 
           mutate(report_date = if_else(is.na(report_date), as.Date(Lab_resdate), as.Date(report_date) )) %>% 
           mutate(report_date = if_else(is.na(report_date), as.Date(consultation_dateHF), as.Date(report_date) )) %>% 
           mutate(report_date = if_else(is.na(report_date), as.Date(Lab_datetaken), as.Date(report_date) )) %>% 
           mutate(report_date = if_else(is.na(report_date), as.Date(patcourse_dateonset), as.Date(report_date) )) %>% 
           mutate(report_date = if_else(is.na(report_date), as.Date(patcourse_datedeath), as.Date(report_date) )) %>% 
           mutate(report_date = if_else(is.na(report_date), as.Date(patcourse_datedischarge), as.Date(report_date) ))%>%
           mutate(patcourse_status = if_else((patcourse_status=='ALIVE' | patcourse_status== 'RECOVERED') & !is.na(patcourse_datedeath), 'DEAD', patcourse_status)) %>%
           mutate(patcourse_status = if_else(is.na(patcourse_status) & !is.na(patcourse_datedeath), 'DEAD', patcourse_status)) %>%
           mutate(patcourse_status = if_else(is.na(patcourse_status) & !is.na(patcourse_datedischarge), 'RECOVERED', patcourse_status)) %>%
           mutate(report_classif = if_else(is.na(report_classif) & (sum(is.na(report_classif))==nrow(df)), 'CONFIRMED', report_classif)) %>% 
           mutate(death = if_else(patcourse_status=='DEAD', 1, 0))%>% 
           mutate(
             age_cat = cut(patinfo_ageonset_years, c(-1,17, 64, 150), labels=c("children", "adults", 'old'))) %>% 
           filter(report_date <=ymd("2020-09-01") & report_date >=ymd("2020-04-01"))%>%
           filter(report_classif=='CONFIRMED') %>% 
           filter(!is.na(patinfo_ageonset_years)) %>% filter(!is.na(patinfo_sex))) %>%
    select(country, report_date, patinfo_sex, patinfo_ageonset_years, death, age_cat, report_classif)
}

lines <- list()
countries_csv <- list.files("~/publication-children/data")
countries_names <- str_trim(str_remove(countries_csv, "_clean.csv"), "both")

df_com = NULL
#for (country in countries_names[grepl('bur', countries_names)]){
for (country in countries_names){
  print(country)
  lines[country] <- readLines(paste("~/publication-children/data/",country, "_clean.csv", sep =''), n = 1)
  #lines[country] <- readLines(here("child_pub/data_analy", glue('{country}_clean.csv')), n = 1)
  
  #******* reading the CSVs *********#
  
  if(grepl(',', lines[country])){
    print(country)
    print("1")
    data1 <- as.data.frame(read.csv(paste("~/publication-children/data/", country, "_clean.csv", sep ='')),na.strings =
                             ".") # , separated
    print(country)}
  else {
    print(country)
    print("2")
    data1 <- as.data.frame(read.csv2(paste("~/publication-children/data/", country, "_clean.csv", sep ='')), na.strings =
                             ".") #} # , separated
    print(country)}
  #print()
  
  data = fetching_data_2(data1, country) # only CONFIRMED cases 
  data_f = bind_rows(df_com, data)
  df_com = data_f
} 
data2 = data_f[order(data_f$report_date), ]
daysinrange <- seq(min(na.omit(data2$report_date)),
                   max(na.omit(data2$report_date)),
                   by = "days")

#coveredsundays <- daysinrange[lubridate::wday(daysinrange) == 1]
coveredsundays <- daysinrange
# Identify the epiweek that each sunday belongs to
epiweeks <- lubridate::epiweek(coveredsundays)

# Extract a nicely formatted date for each Epiweek sunday start
epiweeksundaydate <- format.Date(coveredsundays, format = "%b %d")
epiweekanddate <- paste("EW", epiweeks, "\n", "(",epiweeksundaydate, ")", sep = "")
# epiweekanddate is our final output
# check that all is well with : as.data.frame(coveredsundays,epiweekanddate)
#data2 = data2 %>% 
# mutate(EW=epiweeks)

coveredsundays <- rev(coveredsundays)
epiweekanddate <- rev(epiweekanddate)

coveredsundays2 <- coveredsundays[c(T,F)]
epiweekanddate2 <- epiweekanddate[c(T,F)]

coveredsundays4 <- coveredsundays[c(T,F,F,F)]
epiweekanddate4 <- epiweekanddate[c(T,F,F,F)]

coveredsundays5 <- coveredsundays[c(T,F,F,F,F)]
epiweekanddate5 <- epiweekanddate[c(T,F,F,F,F)]

coveredsundays6 <- coveredsundays[c(T,F,F,F,F,F)]
epiweekanddate6 <- epiweekanddate[c(T,F,F,F,F,F)]

coveredsundays9 <- coveredsundays[c(T,F,F,F,F,F,F,F,F)]
epiweekanddate9 <- epiweekanddate[c(T,F,F,F,F,F,F,F,F)]

age_pop4 = sum(age_pop1$X18.64)
age_pop = sum(age_pop1$total_c) 
pop4 = (100000)/(age_pop4) # 18-64, adult pop
pop = (100000)/(age_pop) # '0-17', children pop
## epi_curves for different age-categories:

# 1. for children 
epi_per_age = data2 %>% group_by(age_cat, report_date) %>% summarise(Reported_this_day = n()) %>%
  filter(age_cat=='children') %>% complete(report_date = seq(as.Date('2020-04-01'), as.Date('2020-09-01') , by = "day")) %>%
  mutate(Reported_this_day = ifelse(is.na(Reported_this_day), 0, Reported_this_day)) %>%
  mutate(Reported_this_day_scaled = Reported_this_day*pop) %>% # cases, scaled by pop
  mutate(cumulat_cases = cumsum(Reported_this_day)) %>% # cumulative cases
  mutate(cumulat_cases_scaled = cumulat_cases*pop) %>% #cumulat cases, scaled by pop
  mutate(sevendayavg_cum=rollmean(x = cumulat_cases_scaled, k=7, align = "right",
                                  fill = NA )) %>% # 7-days rolling average of cum. scaled-cases
  mutate(sevendayavg=rollmean(x = Reported_this_day_scaled, k=7, align = "right",
                              fill = NA )) # 7-days rolling average of scaled-cases

# 2. for adults
epi_per_age4 = data2 %>% group_by(age_cat, report_date) %>% summarise(Reported_this_day = n()) %>%
  filter(age_cat=='adults') %>% complete(report_date = seq(as.Date('2020-04-01'), as.Date('2020-09-01') , by = "day")) %>%
  mutate(Reported_this_day = ifelse(is.na(Reported_this_day), 0, Reported_this_day)) %>%
  mutate(Reported_this_day_scaled4 = Reported_this_day*pop4) %>% # cases, scaled by pop
  mutate(cumulat_cases4 = cumsum(Reported_this_day)) %>% # cumulative cases
  mutate(cumulat_cases_scaled4 = (cumulat_cases4)*pop4) %>% #cumulat cases, scaled by pop
  mutate(sevendayavg_cum4=rollmean(x = cumulat_cases_scaled4, k=7, align = "right",
                                   fill = NA )) %>% # 7-days rolling average of cum. scaled-cases
  mutate(sevendayavg4=rollmean(x = Reported_this_day_scaled4, k=7, align = "right",
                               fill = NA )) # 7-days rolling average of scaled-cases

epi_adul = epi_per_age4 %>%
  mutate(EWs = 
           as.character(epiweek(report_date))) 

epi_adult_group = epi_adul %>% 
  select(age_cat, Reported_this_day, Reported_this_day_scaled4, sevendayavg4, EWs) %>% 
  group_by(EWs) %>% 
  summarise('Reported_casesAVG_epiweek' = 
              mean(Reported_this_day, na.rm=T), 
            'Reported_casesAVG_scaled_epiweek' = 
              mean(Reported_this_day_scaled4, na.rm=T),
            '7daysAVG_epiweek' = 
              mean(sevendayavg4, na.rm=T))

epi_child = epi_per_age %>%
  mutate(EWs = as.character(epiweek(report_date))) 

epi_child_group = epi_child %>% 
  select(age_cat, Reported_this_day, Reported_this_day_scaled, sevendayavg, EWs) %>% 
  group_by(EWs) %>% 
  summarise('Reported_casesAVG_epiweek' = 
              mean(Reported_this_day, na.rm=T), 
            'Reported_casesAVG_scaled_epiweek' = 
              mean(Reported_this_day_scaled, na.rm=T),
            '7daysAVG_epiweek' = 
              mean(sevendayavg, na.rm=T))

#write.csv(epi_child_group, 'children_grouping.csv')
#write.csv(epi_adult_group, 'adults_grouping.csv')

age_com_2 = epi_per_age %>% ggplot() +
  geom_line(aes(x=report_date, y=(sevendayavg), col = "blue"), size = 0.8) +
  geom_line(aes(x=report_date, y=epi_per_age4$sevendayavg4, col = 'red'), size = 0.8, linetype = 'dashed')+
  labs(x=" ", y = " ") + 
  theme(axis.text.x=element_text(size=0.001, angle = 90)) +
  scale_color_identity(name = "Age groups",
                       breaks = c("blue", "red"),
                       labels = c("Children", "Adults"),
                       guide = "legend") +
  scale_x_date(breaks = coveredsundays6,
               labels = epiweekanddate6) +
  theme_bw() +
  theme(legend.position = "top",
        legend.title = element_blank(),
        legend.text = element_text(size = 10))
age_com_2
#write.csv(epi_per_age, 'children_cases.csv') 
#write.csv(epi_per_age4, 'adults_cases.csv')

```
