---
title: "First wave of the COVID-19 pandemic in African island nations: A descriptive study of the variation in the disease burden"
author: "Jyoti Dalal, jyotidalal20@gmail.com"
output: html_document
---

In this project, we used the publicly available COVID-19 data on daily cases and deaths, from the WHO dashboard. Data file could also be downloaded from https://github.com/JyotiDalal93/COVID-19_SSA/blob/main/data_6_country.csv

For more information about this project: https://osf.io/xmqac/

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE)
```

```{r, include=FALSE}
library(tidyverse)
library(zoo)
library(ggrepel)
library(stringi)
library(pander)

my_palette <- c("#56bfa3", "#f79f57" ,"#7570B3","#56B4E9",  #greenblue #lightorange #purplepastel  #lightblue
                "#3758a6" , "#CC79A7" , "#91142c", "#7abcd6", #darkblue #pinkpurple #wine #teal
                "#a3b0c4", "#870476", "#479444", "#3cd6d6" ) # grey, # royalpurple #darkgreen # cyan
# theme
my_theme <- theme_classic() + theme(text = element_text(color = "gray20", size = 12.5),
        rect = element_blank(), # transparent background
        plot.title = (element_text(face = "bold", hjust = 0.5, size = 11)),
        plot.subtitle = (element_text( hjust = 0.5, size = 8, color = alpha("black", 0.7))),
        plot.caption = element_text(size = 6.5, color = "gray50", hjust = 1),
        legend.position = "bottom",
        legend.text = element_text(size = 8),
        legend.key.size = unit(1,"line"),
        axis.title = element_text( color = alpha("black", 0.8), size = 9),
        axis.text = element_text(color = alpha("gray20", 0.7), size = 7 ),
        axis.line = element_line(color = "gray40", size = 0.5),
        axis.line.y = element_blank(),
        panel.grid.major = element_line(color = alpha("gray50", 0.1), size = 0.25),
        strip.background = element_rect(fill=alpha("#0070c0", 1), color = "transparent"),
        strip.text = element_text(face = "bold", colour = 'white'),
        panel.spacing.x = unit(3, "mm"),
        title = element_text(face = "bold"))

theme_set(my_theme)
```


```{r, include=FALSE, warning=FALSE}
lastdate='2020-11-29'

LL_raw <- as.data.frame(read.csv(paste("~/region-epireport/data/",lastdate,"/ConfirmedCases.csv",sep=""),stringsAsFactors = F, na.strings = c("", " ")))

country_info <- read.csv("~/region-epireport/info.csv")

LL_raw = LL_raw %>% 
  mutate(Country = as.character(Country)) %>% 
  mutate(Country = case_when( Country == "Central African Republic" ~ "CAF",
                              Country == "C?te d'Ivoire" ~ "Cote d'Ivoire",
                              Country == "C?te d'Ivoire" ~ "Cote d'Ivoire", 
                              Country == "Democratic Republic of the Congo" ~ "DRC",
                              Country == "Congo (Republic of)" ~ "Congo (Rep)",
                              Country == "Equatorial Guinea" ~ "Eq. Guinea",
                              Country == "Sao Tome and Principe" ~ "Sao Tome e P",
                              Country == "United Republic of Tanzania" ~ "Tanzania",
                              TRUE ~ Country ))

# replace spaces with period in all column names
names(LL_raw) <- str_replace_all(names(LL_raw), " ", ".")

#############################################

# Store as LL_format to fix a bunch of issues
LL_format<-LL_raw  

LL_format$Outcome[which(LL_format$Outcome=="dead")]<-"Dead"
LL_format$Outcome[which(LL_format$Outcome=="recovered")]<-"Recovered"

# impute dates of death
LL_format <- LL_format %>% 
  mutate(Date.of.Death = 
           case_when( is.na(Outcome) ~ "NA", # if no OUTCOME, Date of Death should be NA
                      # if dead and no recorded date of death, 
                      Outcome == "Dead" & is.na(Date.of.Death) ~ 
                        # plug in their date of discharge, or, if  that is missing, the report date
                        ifelse(!is.na(Date.of.Discharge), as.character(Date.of.Discharge) ,  as.character(Reporting_Date)),
                      TRUE ~ as.character(Date.of.Death) ))
# note that the case_when function needs to have all assignments as one type. Here, character

# Recode Dates of Discharge
# As above.
LL_format <- LL_format %>% 
  mutate(Date.of.Discharge = case_when( is.na(Outcome) ~ "NA",
                                        Outcome == "Recovered" & is.na(Date.of.Discharge) ~ 
                                          ifelse(!is.na(Date.of.Discharge), as.character(Date.of.Death) , as.character(Reporting_Date)),
                                        TRUE ~ as.character(Date.of.Discharge) ))

# remove death and recovered dates from individuals neither dead nor recovered
LL_format$Date.of.Death[which(LL_format$Outcome!="Dead")]<-NA
LL_format$Date.of.Discharge[which(LL_format$Outcome!="Recovered")]<-NA

# format dates back to dates.
LL_format$Reporting_Date<-as.Date(LL_format$Reporting_Date)
LL_format$Date.of.Death<-as.Date(LL_format$Date.of.Death)
LL_format$Date.of.Discharge<-as.Date(LL_format$Date.of.Discharge)

# Clip data to last date requested
LL_format<-LL_format[which(LL_format$Reporting_Date<=lastdate),]
LL_format$Outcome[which(LL_format$Outcome=="Dead" & LL_format$Date.of.Death>lastdate)]<-"Alive"
LL_format$Date.of.Death[which(LL_format$Date.of.Death>lastdate)]<-NA
LL_format$Outcome[which(LL_format$Outcome=="Recovered" & LL_format$Date.of.Discharge>lastdate)]<-"Alive"
LL_format$Date.of.Discharge[which(LL_format$Date.of.Discharge>lastdate)]<-NA

# replicate rows based on NumCases, convert to normal LL
LL_format <- LL_format[rep(seq_len(nrow(LL_format)), LL_format$NumCases),]
LL_format$NumCases <- 1

# Back to LL
LL <- LL_format

LL = LL %>% 
  filter(!is.na(FinalEpiClassification))

# Expand dates per country
alldatespercountry <- LL %>% 
  # complete to include all dates, even whe case was not recorded
  complete(Reporting_Date = seq.Date(min(Reporting_Date), max(Reporting_Date), by = "day")) %>%
  # expand to include all dates PER country 
  tidyr::expand (Reporting_Date, Country) %>% 
  # arrange to make things clearer
  arrange(Country, Reporting_Date) %>% 
  # faux variables for death and discharge.
  # needed to bind with data frames later
  mutate(Date.of.Death = Reporting_Date,
         Date.of.Discharge = Reporting_Date)


# Calculate epicurve for cases
casespercountry <-  LL %>%  
  bind_rows(alldatespercountry) %>%
  select(Reporting_Date, Country, NumCases ) %>%
  group_by(Country, Reporting_Date) %>%
  summarise(Reported_this_day = sum(NumCases, na.rm = T))

# Calculate epicurve for deaths as above
deathspercountry <- LL %>% 
  bind_rows(alldatespercountry) %>% 
  select(Date.of.Death, Country, NumCases ) %>%
  group_by(Country, Date.of.Death) %>%
  summarise(Deaths_this_day = sum(NumCases, na.rm = T)) %>% 
  mutate(Reporting_Date = Date.of.Death)

# Calculate epicurve for discharges as above
dischargespercountry <- LL %>% 
  bind_rows(alldatespercountry) %>% 
  select(Date.of.Discharge, Country, NumCases ) %>%
  group_by(Country, Date.of.Discharge) %>%
  summarise(Discharges_this_day = sum(NumCases, na.rm = T)) %>% 
  mutate(Reporting_Date = Date.of.Discharge)


# Combine into single epicurve
Epi_per_country <- inner_join(casespercountry, deathspercountry,
                              by= c("Reporting_Date", "Country"))  %>% 
  inner_join(dischargespercountry,
             by= c("Reporting_Date", "Country")) %>% 
  # sort by date within each country 
  arrange(Country, Reporting_Date) %>% 
  # delete useless vars
  select(Reporting_Date, Country, Reported_this_day, Deaths_this_day, Discharges_this_day ) %>% 
  dplyr::rename(Cases_this_day = Reported_this_day ) %>% 
  # add in EpiWeek 
  mutate(Epiweek = lubridate::epiweek(Reporting_Date)) %>% 
  # Calculate cumulative totals
  group_by(Country) %>% 
  mutate(Cum_cases = cumsum(Cases_this_day),
         Cum_deaths = cumsum(Deaths_this_day),
         Cum_discharges = cumsum(Discharges_this_day),
         Active_cases = Cum_cases - Cum_deaths - Cum_discharges) %>% 
  # remove NA countries (artifact of expansion I think)
  filter(!is.na(Country)) %>% 
  # Calculate crude CFR at each time point
  group_by(Country) %>% 
  mutate(CFR = round(100 * Cum_deaths/Cum_cases, digits = 2 ) ) %>% 
  # Calculate rolling 7 day sums ( past week of cases and deaths)
  mutate(Cases_past_week=rollsum(x = Cases_this_day, k=7, align = "right",  
                                 fill = na.fill(Cases_this_day, 0) ) ,
         Deaths_past_week=rollsum(x = Deaths_this_day, k=7, align = "right",  
                                  fill = na.fill(Deaths_this_day, 0) ) ) %>% 
  left_join(country_info, by = "Country") %>% 
  mutate( Cases_per_million = round((Cum_cases/Population) * 10E5, digits=2) ,
          Deaths_per_million = round((Cum_deaths/Population) * 10E5, digits = 2)) %>% 
  # daily cases and deaths per million
  mutate( Cases_per_million_daily = round((Cases_this_day/Population) * 10E5, digits = 2) ,
          Deaths_per_million_daily = round((Deaths_this_day/Population) * 10E5, digits = 2) ) %>% 
  # daily cases and deaths per million smoothed
  mutate( Cases_per_million_daily_smooth = rollmean(Cases_per_million_daily, k = 7, fill = na.fill(Cases_per_million_daily, "extend")  ) ,
          Deaths_per_million_daily_smooth = rollmean(Deaths_per_million_daily, k = 7, fill = na.fill(Deaths_per_million_daily, "extend") )) %>% 
  ungroup() 

daysinrange <- seq(min(na.omit(LL$Reporting_Date)),max(na.omit(LL$Reporting_Date)),
                   by = "days")
# Subset just the Sundays. Epiweeks start on Sunday, so we will need these to assign labels
coveredsundays <- daysinrange[lubridate::wday(daysinrange) == 1]
# Identify the epiweek that each sunday belongs to
epiweeks <- lubridate::epiweek(coveredsundays) 
# Extract a nicely formatted date for each Epiweek sunday start
epiweeksundaydate <- format.Date(coveredsundays, format = "%b %d")
epiweekanddate <-  paste0(epiweeksundaydate, "\n", "(", "EW", epiweeks,")" )
# epiweekanddate is our final output
# check that all is well with : as.data.frame(coveredsundays,epiweekanddate)

coveredsundays <- rev(coveredsundays)
epiweekanddate <- rev(epiweekanddate)
# every other sunday. For tight plots
coveredsundays2 <- coveredsundays[c(T,F)] 
epiweekanddate2 <- epiweekanddate[c(T,F)]
# every third sunday. For really tight plots
coveredsundays3 <- coveredsundays[c(T,F,F)] 
epiweekanddate3 <- epiweekanddate[c(T,F,F)]
# every fourth sunday. For really really tight plots
coveredsundays4 <- coveredsundays[c(T,F,F,F)] 
epiweekanddate4 <- epiweekanddate[c(T,F,F,F)]
coveredsundays5 <- coveredsundays[c(T,F,F,F,F)] 
epiweekanddate5 <- epiweekanddate[c(T,F,F,F,F)]
coveredsundays6 <- coveredsundays[c(T,F,F,F,F,F)] 
epiweekanddate6 <- epiweekanddate[c(T,F,F,F,F,F)]

################################################
#  OVERVIEW TABLES
################################################

all_country_tab <- Epi_per_country %>% 
  group_by(Country) %>% 
  slice(which.max(Reporting_Date)) %>% 
  dplyr::rename(Cases = Cum_cases, Deaths = Cum_deaths ) %>% 
  mutate(`Crude \nCFR (%)` = round(100 * Deaths/Cases, digits = 2 ) ) %>% 
  mutate( `Cases per million` = (Cases/Population) * 10E5 ,
          `Deaths per million` = (Deaths/Population) * 10E5 ) %>% 
  select(Country, Cases,`Cases per million`, Deaths, `Deaths per million`) %>% 
  arrange(-Cases) %>% 
  left_join(country_info) %>% 
  ungroup()

regional_tab <- Epi_per_country %>% 
  select(Reporting_Date, Country, Cum_cases, Cum_deaths, Region, Population) %>% 
  group_by(Country) %>% 
  slice(which.max(Reporting_Date)) %>% 
  group_by(Region) %>% 
  mutate(Cum_cases_region = sum(Cum_cases),
         Cum_deaths_region = sum(Cum_deaths),
         Population = sum(Population)) %>% 
  slice(which.max(Reporting_Date)) %>% 
  rename(Cases = Cum_cases_region, Deaths= Cum_deaths_region) %>% 
  mutate(`Crude \nCFR (%)` = round(100 * Deaths/Cases, digits = 2)) %>% 
  mutate( `Cases per million` = (Cases/Population) * 10E5 ,
          `Deaths per million` = (Deaths/Population) * 10E5 ) %>% 
  select(Region, Cases,`Cases per million`, Deaths, `Deaths per million`) %>% 
  arrange(-Cases) %>% 
  ungroup()

```

#### Figure 1: Daily number of cases in each island nation ####

```{r fig.width=5.5, fig.height=3.5, fig.align='center'}
my_country = "Cabo Verde" # 1

Epi_per_country %>% 
  filter(Country == "Cabo Verde" ) %>% 
  select(Reporting_Date, Cases_this_day) %>% 
  mutate(sevendayavg=rollmean(x = Cases_this_day, k=7, align = "right",  
                              fill = na.fill(Cases_this_day, 0) )) %>% 
  ggplot() +
  geom_bar( aes(x= Reporting_Date, y= Cases_this_day), stat = "identity", 
            fill = "green", colour = alpha("white", 0.75)) +
  geom_line(aes(x=Reporting_Date, y=sevendayavg, lty = "7-day rolling average")) +
  scale_x_date(breaks = coveredsundays4,
               labels = epiweekanddate4,
               limits = c(as.Date("2020-03-01"), NA ))+
  labs(x="Date of Reporting", y =paste("Daily COVID-19 incidence, ", my_country)) +
  theme(legend.title = element_blank())
```

```{r fig.width=5.5, fig.height=3.5, fig.align='center'}
my_country = "Comoros" # 2

Epi_per_country %>% 
  filter(Country == my_country ) %>% 
  select(Reporting_Date, Cases_this_day) %>% 
  mutate(sevendayavg=rollmean(x = Cases_this_day, k=7, align = "right",  
                              fill = na.fill(Cases_this_day, 0) )) %>% 
  ggplot() +
  geom_bar( aes(x= Reporting_Date, y= Cases_this_day), stat = "identity", 
            fill = "blue", colour = alpha("white", 0.75)) +
  geom_line(aes(x=Reporting_Date, y=sevendayavg, lty = "7-day rolling average")) +
  scale_x_date(breaks = coveredsundays4,
               labels = epiweekanddate4,
               limits = c(as.Date("2020-03-01"), NA ))+
  labs(x="Date of Reporting", y =paste("Daily COVID-19 incidence, ", my_country)) +
  theme(legend.title = element_blank())
```

```{r fig.width=5.5, fig.height=3.5, fig.align='center'}
my_country = "Madagascar"  # 3

Epi_per_country %>% 
  filter(Country == my_country ) %>% 
  select(Reporting_Date, Cases_this_day) %>% 
  mutate(sevendayavg=rollmean(x = Cases_this_day, k=7, align = "right",  
                              fill = na.fill(Cases_this_day, 0) )) %>% 
  ggplot() +
  geom_bar( aes(x= Reporting_Date, y= Cases_this_day), stat = "identity", 
            fill = "blue", colour = alpha("white", 0.75)) +
  geom_line(aes(x=Reporting_Date, y=sevendayavg, lty = "7-day rolling average")) +
  scale_x_date(breaks = coveredsundays4,
               labels = epiweekanddate4,
               limits = c(as.Date("2020-03-01"), NA ))+
  labs(x="Date of Reporting", y =paste("Daily COVID-19 incidence, ", my_country)) +
  theme(legend.title = element_blank())
```


```{r fig.width=5.5, fig.height=3.5, fig.align='center'}
my_country = "Mauritius" #4

Epi_per_country %>% 
  filter(Country == my_country ) %>% 
  select(Reporting_Date, Cases_this_day) %>% 
  mutate(sevendayavg=rollmean(x = Cases_this_day, k=7, align = "right",  
                              fill = na.fill(Cases_this_day, 0) )) %>% 
  ggplot() +
  geom_bar( aes(x= Reporting_Date, y= Cases_this_day), stat = "identity", 
            fill = "darkolivegreen2", colour = alpha("white", 0.75)) +
  geom_line(aes(x=Reporting_Date, y=sevendayavg, lty = "7-day rolling average")) +
  scale_x_date(breaks = coveredsundays4,
               labels = epiweekanddate4,
               limits = c(as.Date("2020-03-01"), NA ))+
  labs(x="Date of Reporting", y =paste("Daily COVID-19 incidence, ", my_country)) +
  theme(legend.title = element_blank())
```

```{r fig.width=5.5, fig.height=3.5, fig.align='center'}
my_country = "Sao Tome e P" #5

Epi_per_country %>% 
  filter(Country == my_country ) %>% 
  select(Reporting_Date, Cases_this_day) %>% 
  mutate(sevendayavg=rollmean(x = Cases_this_day, k=7, align = "right",  
                              fill = na.fill(Cases_this_day, 0) )) %>% 
  ggplot() +
  geom_bar( aes(x= Reporting_Date, y= Cases_this_day), stat = "identity", 
            fill = "#D95F02", colour = alpha("white", 0.75)) +
  geom_line(aes(x=Reporting_Date, y=sevendayavg, lty = "7-day rolling average")) +
  scale_x_date(breaks = coveredsundays4,
               labels = epiweekanddate4,
               limits = c(as.Date("2020-03-01"), NA ))+
  labs(x="Date of Reporting", y =paste("Daily COVID-19 incidence, ", my_country)) +
  theme(legend.title = element_blank())
```


```{r fig.width=5.5, fig.height=3.5, fig.align='center'}
my_country = "Seychelles" # 6

Epi_per_country %>% 
  filter(Country == my_country ) %>% 
  select(Reporting_Date, Cases_this_day) %>% 
  mutate(sevendayavg=rollmean(x = Cases_this_day, k=7, align = "right",  
                              fill = na.fill(Cases_this_day, 0) )) %>% 
  ggplot() +
  geom_bar( aes(x= Reporting_Date, y= Cases_this_day), stat = "identity", 
            fill = "pink", colour = alpha("white", 0.75)) +
  geom_line(aes(x=Reporting_Date, y=sevendayavg, lty = "7-day rolling average")) +
  scale_x_date(breaks = coveredsundays4,
               labels = epiweekanddate4,
               limits = c(as.Date("2020-03-01"), NA ))+
  labs(x="Date of Reporting", y =paste("Daily COVID-19 incidence, ", my_country)) +
  theme(legend.title = element_blank())
```

#### Table 2: Cases, deaths, CFRs, cases per million and deaths per million for all the island nations are shown below: ####

```{r}

countries_to_label <- c("Cabo Verde", "Comoros", "Madagascar", "Mauritius", "Sao Tome e P", "Seychelles")

regional_tab <- Epi_per_country %>% 
  select(Reporting_Date, Country, Cum_cases, Cum_deaths, Region, Population) %>% 
  group_by(Country) %>% 
  slice(which.max(Reporting_Date)) %>% 
  group_by(Region) %>% 
  mutate(Cum_cases_region = sum(Cum_cases),
         Cum_deaths_region = sum(Cum_deaths),
         Population = sum(Population)) %>% 
  slice(which.max(Reporting_Date)) %>% 
  rename(Cases = Cum_cases_region, Deaths= Cum_deaths_region) %>% 
  mutate(`Crude \nCFR (%)` = round(100 * Deaths/Cases, digits = 2 ) ) %>% 
  mutate( `Cases per million` = (Cases/Population) * 10E5 ,
          `Deaths per million` = (Deaths/Population) * 10E5 ) %>% 
  select(Region, Cases,`Cases per million`, Deaths, `Deaths per million`) %>% 
  arrange(-Cases) %>% 
  ungroup()

country_tab_myreg <- all_country_tab %>%
  mutate(`Crude CFR` = round((100* Deaths/Cases), digits = 2 ),
         `CasesPerMil` = round(`Cases per million`, 2 ),
         `Deaths per million` = round(`Deaths per million`, 2 )) %>% 
  left_join(country_info) %>% 
  filter(Country %in% countries_to_label) %>% 
  select(-c(iso_code, Region) ) %>% 
  select(Country, Cases, Deaths, `Crude CFR`, Population, 
         `Cases per million`, `Deaths per million` ) %>% 
  as_tibble()

panderOptions('table.continues', '')
panderOptions('table.caption.prefix', '')
panderOptions('table.style', 'rmarkdown')
pander(country_tab_myreg, split.table = 1000, split.cells = 8)
```

### Supplementary Information ##

#### 1. Figures S1-S2 showing cases per million and deaths per million among different island nations ####

```{r fig.width=5.5, fig.height=3.5, fig.align='center'}
countries_to_label <- c("Cabo Verde", "Comoros", "Madagascar", "Mauritius", "Sao Tome e P", "Seychelles")

country_comparisons <- Epi_per_country %>% 
  filter(Country %in% countries_to_label)  %>% 
  # paste region and cases at maxdate. Needed for plot
  group_by(Country) %>% 
  mutate(label_max_cases = if_else(Reporting_Date == max(Reporting_Date, na.rm=T),
                                   paste(as.character(Country), "\n",  
                                         max(Cases_per_million, na.rm = T)) ,
                                   NA_character_),
         label_max_deaths = if_else(Reporting_Date == max(Reporting_Date, na.rm=T),
                                    paste(as.character(Country), "\n",  
                                          max(Deaths_per_million, na.rm = T)) ,
                                    NA_character_))

####
Epi_per_country %>% 
  filter(Country == "Cabo Verde")%>% 
  select(Reporting_Date, Cases_per_million, Country) %>% 
  ggplot() +
  geom_line( aes(x= Reporting_Date, y= Cases_per_million, group = Country),
             colour = alpha("black",0.6 ), size = 0.4) +
  geom_line(data= country_comparisons,
            aes(x= Reporting_Date, y= Cases_per_million, colour = Country),
            size = 1) +
  geom_label_repel(data= country_comparisons,
                   aes( x= Reporting_Date, y= Cases_per_million, 
                        colour = Country, label = label_max_cases),
                   na.rm = TRUE, size = 2.8, fill = alpha("white", 0.8),
                   fontface = "bold") +
  scale_color_manual(values= my_palette) +
  #ggtitle(paste("Cumulative cases per million for all African island countries")) +
  scale_x_date(breaks = coveredsundays4,
               labels = epiweekanddate4,
               limits = c(as.Date("2020-03-01"), NA ))+
  labs(x="Date of Reporting", y ="Cases per million") +
  theme(legend.position = "none")
```

```{r fig.width=5.5, fig.height=3.5, fig.align='center'}
#####
Epi_per_country %>% 
  filter(Country == "Seychelles")%>% 
  #filter(Region == my_region )%>% 
  select(Reporting_Date, Deaths_per_million, Country) %>% 
  ggplot() +
  geom_line( aes(x= Reporting_Date, y= Deaths_per_million, group = Country),
             colour = alpha("black",0.6 ), size = 0.4) +
  geom_line(data= country_comparisons,
            aes(x= Reporting_Date, y= Deaths_per_million, colour = Country),
            size = 1) +
  geom_label_repel(data= country_comparisons,
                   aes( x= Reporting_Date, y= Deaths_per_million, 
                        colour = Country, label = label_max_deaths),
                   na.rm = TRUE, size = 2.8, fill = alpha("white", 0.8),
                   fontface = "bold") +
  scale_color_manual(values= my_palette) +
  #ggtitle(paste("Cumulative deaths per million for all island countries in Africa"),
   #       paste(my_country, "and select others highlighted")) +
  scale_x_date(breaks = coveredsundays4,
               labels = epiweekanddate4,
               limits = c(as.Date("2020-03-01"), NA ))+
  labs(x="Date of Death", y ="Deaths per million") +
  theme(legend.position = "none")
```

#### 2. Comparison of island nations with the non-island ones from the same geographical region ####

= Table S2

1. Central Africa
```{r}

my_region = 'Central Africa' 
regional_tab <- Epi_per_country %>% 
  select(Reporting_Date, Country, Cum_cases, Cum_deaths, Region, Population) %>% 
  group_by(Country) %>% 
  slice(which.max(Reporting_Date)) %>% 
  group_by(Region) %>% 
  mutate(Cum_cases_region = sum(Cum_cases),
         Cum_deaths_region = sum(Cum_deaths),
         Population = sum(Population)) %>% 
  slice(which.max(Reporting_Date)) %>% 
  rename(Cases = Cum_cases_region, Deaths= Cum_deaths_region) %>% 
  mutate(`Crude \nCFR (%)` = round(100 * Deaths/Cases, digits = 2 ) ) %>% 
  mutate( `Cases per million` = (Cases/Population) * 10E5 ,
          `Deaths per million` = (Deaths/Population) * 10E5 ) %>% 
  select(Region, Cases,`Cases per million`, Deaths, `Deaths per million`) %>% 
  arrange(-Cases) %>% 
  ungroup()


country_tab_myreg <- all_country_tab %>%
  mutate(`Crude CFR` = round((100* Deaths/Cases), digits = 2 ),
         `CasesPerMil` = round(`Cases per million`, 2 ),
         `Deaths per million` = round(`Deaths per million`, 2 )) %>% 
  left_join(country_info) %>% 
  filter(Region == my_region) %>% 
  select(-c(iso_code, Region) ) %>% 
  select(Country, Cases, Deaths, `Crude CFR`, Population, 
         `Cases per million`, `Deaths per million` ) %>% 
  as_tibble()

panderOptions('table.continues', '')
panderOptions('table.caption.prefix', '')
panderOptions('table.style', 'rmarkdown')

pander(country_tab_myreg, split.table = 1000, split.cells = 8)

```

2. Eastern Africa

```{r}

my_region = 'Eastern Africa' 
regional_tab <- Epi_per_country %>% 
  select(Reporting_Date, Country, Cum_cases, Cum_deaths, Region, Population) %>% 
  group_by(Country) %>% 
  slice(which.max(Reporting_Date)) %>% 
  group_by(Region) %>% 
  mutate(Cum_cases_region = sum(Cum_cases),
         Cum_deaths_region = sum(Cum_deaths),
         Population = sum(Population)) %>% 
  slice(which.max(Reporting_Date)) %>% 
  rename(Cases = Cum_cases_region, Deaths= Cum_deaths_region) %>% 
  mutate(`Crude \nCFR (%)` = round(100 * Deaths/Cases, digits = 2 ) ) %>% 
  mutate( `Cases per million` = (Cases/Population) * 10E5 ,
          `Deaths per million` = (Deaths/Population) * 10E5 ) %>% 
  select(Region, Cases,`Cases per million`, Deaths, `Deaths per million`) %>% 
  arrange(-Cases) %>% 
  ungroup()


country_tab_myreg <- all_country_tab %>%
  mutate(`Crude CFR` = round((100* Deaths/Cases), digits = 2 ),
         `CasesPerMil` = round(`Cases per million`, 2 ),
         `Deaths per million` = round(`Deaths per million`, 2 )) %>% 
  left_join(country_info) %>% 
  filter(Region == my_region) %>% 
  select(-c(iso_code, Region) ) %>% 
  select(Country, Cases, Deaths, `Crude CFR`, Population, 
         `Cases per million`, `Deaths per million` ) %>% 
  as_tibble()

panderOptions('table.continues', '')
panderOptions('table.caption.prefix', '')
panderOptions('table.style', 'rmarkdown')

pander(country_tab_myreg, split.table = 1000, split.cells = 8)

```

3. Western Africa

```{r}
my_region = 'Western Africa' 
regional_tab <- Epi_per_country %>% 
  select(Reporting_Date, Country, Cum_cases, Cum_deaths, Region, Population) %>% 
  group_by(Country) %>% 
  slice(which.max(Reporting_Date)) %>% 
  group_by(Region) %>% 
  mutate(Cum_cases_region = sum(Cum_cases),
         Cum_deaths_region = sum(Cum_deaths),
         Population = sum(Population)) %>% 
  slice(which.max(Reporting_Date)) %>% 
  rename(Cases = Cum_cases_region, Deaths= Cum_deaths_region) %>% 
  mutate(`Crude \nCFR (%)` = round(100 * Deaths/Cases, digits = 2 ) ) %>% 
  mutate( `Cases per million` = (Cases/Population) * 10E5 ,
          `Deaths per million` = (Deaths/Population) * 10E5 ) %>% 
  select(Region, Cases,`Cases per million`, Deaths, `Deaths per million`) %>% 
  arrange(-Cases) %>% 
  ungroup()


country_tab_myreg <- all_country_tab %>%
  mutate(`Crude CFR` = round((100* Deaths/Cases), digits = 2),
         `CasesPerMil` = round(`Cases per million`, 2 ),
         `Deaths per million` = round(`Deaths per million`, 2)) %>% 
  left_join(country_info) %>% 
  filter(Region == my_region) %>% 
  select(-c(iso_code, Region) ) %>% 
  select(Country, Cases, Deaths, `Crude CFR`, Population, 
         `Cases per million`, `Deaths per million` ) %>% 
  as_tibble()

panderOptions('table.continues', '')
panderOptions('table.caption.prefix', '')
panderOptions('table.style', 'rmarkdown')

pander(country_tab_myreg, split.table = 1000, split.cells = 8)

```

4. Southern Africa

```{r}
my_region = 'Southern Africa' 
regional_tab <- Epi_per_country %>% 
  select(Reporting_Date, Country, Cum_cases, Cum_deaths, Region, Population) %>% 
  group_by(Country) %>% 
  slice(which.max(Reporting_Date)) %>% 
  group_by(Region) %>% 
  mutate(Cum_cases_region = sum(Cum_cases),
         Cum_deaths_region = sum(Cum_deaths),
         Population = sum(Population)) %>% 
  slice(which.max(Reporting_Date)) %>% 
  rename(Cases = Cum_cases_region, Deaths= Cum_deaths_region) %>% 
  mutate(`Crude \nCFR (%)` = round(100 * Deaths/Cases, digits = 2 ) ) %>% 
  mutate( `Cases per million` = (Cases/Population) * 10E5 ,
          `Deaths per million` = (Deaths/Population) * 10E5 ) %>% 
  select(Region, Cases,`Cases per million`, Deaths, `Deaths per million`) %>% 
  arrange(-Cases) %>% 
  ungroup()


country_tab_myreg <- all_country_tab %>%
  mutate(`Crude CFR` = round((100* Deaths/Cases), digits = 2),
         `CasesPerMil` = round(`Cases per million`, 2 ),
         `Deaths per million` = round(`Deaths per million`, 2)) %>% 
  left_join(country_info) %>% 
  filter(Region == my_region) %>% 
  select(-c(iso_code, Region) ) %>% 
  select(Country, Cases, Deaths, `Crude CFR`, Population, 
         `Cases per million`, `Deaths per million` ) %>% 
  as_tibble()

panderOptions('table.continues', '')
panderOptions('table.caption.prefix', '')
panderOptions('table.style', 'rmarkdown')

pander(country_tab_myreg, split.table = 1000, split.cells = 8)

```

5. Northern Africa

```{r}
my_region = 'Northern Africa' 
regional_tab <- Epi_per_country %>% 
  select(Reporting_Date, Country, Cum_cases, Cum_deaths, Region, Population) %>% 
  group_by(Country) %>% 
  slice(which.max(Reporting_Date)) %>% 
  group_by(Region) %>% 
  mutate(Cum_cases_region = sum(Cum_cases),
         Cum_deaths_region = sum(Cum_deaths),
         Population = sum(Population)) %>% 
  slice(which.max(Reporting_Date)) %>% 
  rename(Cases = Cum_cases_region, Deaths= Cum_deaths_region) %>% 
  mutate(`Crude \nCFR (%)` = round(100 * Deaths/Cases, digits = 2 ) ) %>% 
  mutate( `Cases per million` = (Cases/Population) * 10E5 ,
          `Deaths per million` = (Deaths/Population) * 10E5 ) %>% 
  select(Region, Cases,`Cases per million`, Deaths, `Deaths per million`) %>% 
  arrange(-Cases) %>% 
  ungroup()


country_tab_myreg <- all_country_tab %>%
  mutate(`Crude CFR` = round((100* Deaths/Cases), digits = 2),
         `CasesPerMil` = round(`Cases per million`, 2 ),
         `Deaths per million` = round(`Deaths per million`, 2)) %>% 
  left_join(country_info) %>% 
  filter(Region == my_region) %>% 
  select(-c(iso_code, Region) ) %>% 
  select(Country, Cases, Deaths, `Crude CFR`, Population, 
         `Cases per million`, `Deaths per million` ) %>% 
  as_tibble()

panderOptions('table.continues', '')
panderOptions('table.caption.prefix', '')
panderOptions('table.style', 'rmarkdown')

pander(country_tab_myreg, split.table = 1000, split.cells = 8)

```
